Containers and R---the Rockerverse and beyond

### 2020-01-10 manuscript.Rmd
add data science images section

>>>>

## Data Science images

The notion of _Data Science_ is widely discussed among all academic disciplines.
Using computers and complex software to make sense of (large amounts of) collected or simulated data is a cross-cutting phenomenon leading to _X Data Science_ disciplines, seminars, and study majors, there are also Docker images with commonly used tools and software stacks pre-installed that are often used in combination for analysing data.
A large number of bespoke computing environments are created this way.
Here we only mention a few image stacks as examples that are not based on Rocker images but prominently include installations of R.

The [**Jupyter Docker Stacks** project](https://github.com/jupyter/docker-stacks/) are a set of ready-to-run Docker images containing Jupyter applications and interactive computing tools \citep{project_jupyter_jupyter_2018}.
The `jupyter/r-notebook` image includes R and "popular packages", and R is also included in the catchall `datascience-notebook` image\footnote{\href{https://jupyter-docker-stacks.readthedocs.io/en/latest/using/selecting.html}{https://jupyter-docker-stacks.readthedocs.io/en/latest/using/selecting.html}}.
For example, these images allow users to quickly start a Jupyter Notebook server locally or build their own specialised images on top of stable toolsets.

**Kaggle** provides the [`gcr.io/kaggle-images/rstats`](https://hub.docker.com/r/kaggle/rstats) image (previously `kaggle/rstats`) and [corresponding `Dockerfile`](https://github.com/Kaggle/docker-rstats) for usage in their Machine Learning competitions and easy access to the associated datasets.
It includes machine learling libraries such as Tensorflow and Keras, and also configures the \pkg{reticulate} package.
The image uses a base image with _all packages from CRAN_, `gcr.io/kaggle-images/rcran`, which requires a Google Cloud Build as Docker Hub would time out\footnote{Originally, a stacked collection of over 20 images with automated builds on Docker Hub was used, see \href{https://web.archive.org/web/20190606043353/http://blog.kaggle.com/2016/02/05/how-to-get-started-with-data-science-in-containers/}{https://web.archive.org/web/20190606043353/http://blog.kaggle.com/2016/02/05/how-to-get-started-with-data-science-in-containers/} and \href{https://hub.docker.com/r/kaggle/rcran/dockerfile}{https://hub.docker.com/r/kaggle/rcran/dockerfile}}.
The final extracted image size is over 25GB, which makes it debatable if having everything available is actually convenient.

```{bash kaggle_images, eval=FALSE, include=FALSE}
docker pull gcr.io/kaggle-images/rcran
docker pull kaggle/rcran # over 22 months old
docker pull gcr.io/kaggle-images/rstats
docker images | grep kaggle
```

As a further example, [**Radiant** project](https://radiant-rstats.github.io/docs/) provides several images, e.g., [`vnijs/rsm-msba-spark`](https://hub.docker.com/r/vnijs/rsm-msba-spark), for their browser-based business analytics interface based on \pkg{Shiny} (`Dockerfile` [on GitHub](https://github.com/radiant-rstats/docker)) and for use in education as part of an MSc course.
As data science often applies a multitude of tools, this image favours inclusion over selection and features Python, Postgres, JupyterLab and Visual Studio Code besides R and RStudio, bringing the image size up to 9GB.

```{bash radiant_images, eval=FALSE, include=FALSE}
docker pull vnijs/rsm-msba-spark
docker images | grep vnijs
```
